---
title: "R coding Exercise"
output: 
  html_document:
    toc: FALSE
---
### Loading and Checking Data Section
<<<<<<< HEAD

```{r}
#Loads the "dslabs" and "tidyverse" packages
library(dslabs)
library(tidyverse)

#Pulls up the help page for the gapminder data
help(gapminder)
#Pulls up the structure for the gapminder data
str(gapminder)
#Summary of the gapminder data
summary(gapminder)
#Determines the type of object that gapminder is
class(gapminder)

```

### Processing Data Section

```{r}
#using the gapminder data to create a new object that contains only the African countries
africadata <- subset(gapminder, continent == "Africa")

#Creating a new object with africadata that only contains infant_mortality and life_expectancy
africaIMLE <- select(africadata, infant_mortality, life_expectancy)

#Creating a new object with africadata that only contains population and life_expectancy
africaPLE <- select(africadata, population, life_expectancy)

```

### Plotting Section

```{r}
#Plotting life expectancy as a function of infant mortality using the africaIMLE dataframe
ggplot(africaIMLE, aes(x = infant_mortality, y = life_expectancy)) +
      geom_point() +
      xlab("Infant Mortality") +
      ylab("Life Expectancy")
#A negative correlation shows between the two



#Plotting life expectancy as a function of population on a log scale using the africaPLE dataframe
ggplot(africaPLE, aes(x = log(population), y = life_expectancy)) +
      geom_point() +
      xlab("Population (Logarithmic Scale)") +
      ylab("Life Expectancy")
#Streaks of positively-correlated data appear on this graph

```

### More Data Processing Section

The reason for the second graph's streaks of data is because there is also different years shown in the dataset for each country. Over time, population size increases as well as life expectancy.

```{r}
#Code to figure out which years have missing data for infant mortality in the africadata dataset
africaYIM <- select(africadata, year, infant_mortality)
#This creates a new data frame that only contains the year data and infant mortality data

is.IMna <- africaYIM[rowSums(is.na(africaYIM)) > 0, ]
#This creates a new data frame that only contains the year data with NA values for infant mortality


years_with_NA <- unique(is.IMna)
#This creates a new data frame that contains the years that have NA infant mortality data

years_with_NA
```

Based on this output, we can see that for years up to 1981 and then for 2016, there is missing infant mortality data. Next we will extract all the data for only the year 2000 from the africadata object.

```{r}
africa2000 <- filter(africadata, year == "2000")
str(africa2000)
```

### More Plotting Section

This section will graph the same correlations as above, but only for the year 2000.

```{r}
#This graphs life expectancy as a function of infant mortality for the year 2000 in Africa.
ggplot(africa2000, aes(x = infant_mortality, y = life_expectancy)) +
      geom_point() +
      xlab("Infant Mortality") +
      ylab("Life Expectancy")
#There is a negative correlation between the two.

#This graphs life expectancy as a function of population for the year 2000 in Africa.
ggplot(africa2000, aes(x = log(population), y = life_expectancy)) +
      geom_point() +
      xlab("Population (Logarithmic Scale)") +
      ylab("Life Expectancy")
#There is no noticable correlation.

```

### A Simple Fit Section
```{r}
#Fitting a linear model to the 2000 Africa data, with life expectancy as the outcome and infant mortality as the predictor

fit1 <- lm(life_expectancy ~ infant_mortality, data = africa2000)
summary(fit1)
```

Based on the p-value, we can conclude that infant mortality had a statistically significant effect on life expectancy in Africa during the year 2000.

```{r}
#Adding a linear model to the 2000 Africa data, with life expectancy as the outcome and population as the predictor

fit2 <- lm(life_expectancy ~ population, data = africa2000)
summary(fit2)
```

Based on the p-value, we can conclude that population did not have a statistically significant effect on life expectancy in Africa during the year 2000.

$$\\[.25in]$$

#Now lets look at infant mortality and fertility in China. Are they good predictors of population?

#before we start, go ahead and download the "broom" package
```{r}
library(broom)
```

#Create an object that only contains data from China

```{r}
chinaadata <- subset(gapminder, country == "China")
```

#Creating new data framed with our selected variables
```{r}
chinaIMP <- chinadata[c("infant_mortality","population")]
chinaFP <- chinadata[c("fertility","population")]

summary(chinaFP)
summary(chinaIMP)

```

###Plotting

#Plotting infant mortality as a function of population(log scale)
```{r}
ggplot(chinaIMP, aes(x = log(population), y = infant_mortality)) + geom_point()
```

#Plotting fertility as a function of population(log scale)
```{r}
ggplot(chinaFP, aes(x = log(population), y = fertility)) + geom_point()
```
#Both have a negative association, but it appears to be a bit stronger with infant mortality as a function of population
#Note, there are 6 rows that cotain missing values in graph 1 and 1 row missing in graph #2. These should not affect the linear model.

###Creating Linear Models

#Infant mortality vs. population
```{R}
fit3 <- lm(formula = population ~ infant_mortality, data = chinaIMP)

summary(fit3)
```
#Fertility vs. population
```{r}
fit4 <- lm(formula = population ~ fertility, data = chinaFP)

summary(fit4)

```
#Since both have p-values <0.05, both are appropriate predictors of population

#With the help of the broom package, we can look at this in a graph form
```{r}
tidy(fit3)
tidy(fit3)
```
=======

```{r}
#Loads the "dslabs" and "tidyverse" packages
library(dslabs)
library(tidyverse)

#Pulls up the help page for the gapminder data
help(gapminder)
#Pulls up the structure for the gapminder data
str(gapminder)
#Summary of the gapminder data
summary(gapminder)
#Determines the type of object that gapminder is
class(gapminder)

```

### Processing Data Section

```{r}
#using the gapminder data to create a new object that contains only the African countries
africadata <- subset(gapminder, continent == "Africa")

#Creating a new object with africadata that only contains infant_mortality and life_expectancy
africaIMLE <- select(africadata, infant_mortality, life_expectancy)

#Creating a new object with africadata that only contains population and life_expectancy
africaPLE <- select(africadata, population, life_expectancy)

```

### Plotting Section

```{r}
#Plotting life expectancy as a function of infant mortality using the africaIMLE dataframe
ggplot(africaIMLE, aes(x = infant_mortality, y = life_expectancy)) +
      geom_point() +
      xlab("Infant Mortality") +
      ylab("Life Expectancy")
#A negative correlation shows between the two



#Plotting life expectancy as a function of population on a log scale using the africaPLE dataframe
ggplot(africaPLE, aes(x = log(population), y = life_expectancy)) +
      geom_point() +
      xlab("Population (Logarithmic Scale)") +
      ylab("Life Expectancy")
#Streaks of positively-correlated data appear on this graph

```

### More Data Processing Section

The reason for the second graph's streaks of data is because there is also different years shown in the dataset for each country. Over time, population size increases as well as life expectancy.

```{r}
#Code to figure out which years have missing data for infant mortality in the africadata dataset
africaYIM <- select(africadata, year, infant_mortality)
#This creates a new data frame that only contains the year data and infant mortality data

is.IMna <- africaYIM[rowSums(is.na(africaYIM)) > 0, ]
#This creates a new data frame that only contains the year data with NA values for infant mortality


years_with_NA <- unique(is.IMna)
#This creates a new data frame that contains the years that have NA infant mortality data

years_with_NA
```

Based on this output, we can see that for years up to 1981 and then for 2016, there is missing infant mortality data. Next we will extract all the data for only the year 2000 from the africadata object.

```{r}
africa2000 <- filter(africadata, year == "2000")
str(africa2000)
```

### More Plotting Section

This section will graph the same correlations as above, but only for the year 2000.

```{r}
#This graphs life expectancy as a function of infant mortality for the year 2000 in Africa.
ggplot(africa2000, aes(x = infant_mortality, y = life_expectancy)) +
      geom_point() +
      xlab("Infant Mortality") +
      ylab("Life Expectancy")
#There is a negative correlation between the two.

#This graphs life expectancy as a function of population for the year 2000 in Africa.
ggplot(africa2000, aes(x = log(population), y = life_expectancy)) +
      geom_point() +
      xlab("Population (Logarithmic Scale)") +
      ylab("Life Expectancy")
#There is no noticable correlation.

```

### A Simple Fit Section
```{r}
#Fitting a linear model to the 2000 Africa data, with life expectancy as the outcome and infant mortality as the predictor

fit1 <- lm(life_expectancy ~ infant_mortality, data = africa2000)
summary(fit1)
```

Based on the p-value, we can conclude that infant mortality had a statistically significant effect on life expectancy in Africa during the year 2000.

```{r}
#Adding a linear model to the 2000 Africa data, with life expectancy as the outcome and population as the predictor

fit2 <- lm(life_expectancy ~ population, data = africa2000)
summary(fit2)
```

Based on the p-value, we can conclude that population did not have a statistically significant effect on life expectancy in Africa during the year 2000.


>>>>>>> main


